---
title: "Introduction to Quantitative Finance for R: Financial Networks"
author: "Andrea Bernard, Scott Jacobs, Justin Salvador, Haixia Zhang"
date: Feb 6, 2016
output: ioslides_presentation
---

##Presentation Objectives

We are presenting this information to you so that you may

> - Have a fundamental understanding of Financial Networks and their properties.

> - Identify Systemically Important Financial Institutions.

> - Add Financial Network Analysis to your investment analysis framework.

## Obj 1: Understanding Financial Networks Informs Trading

Financial networks are a map of which bank has lent money to some other bank. Analyzing the network can help identify banks that pose inherent risks to the system.

> - Guide overall portfolio strategy
> - Inform clients of potential pitfalls
> - Identify specific bank stocks to short

## Properties of Financial Networks 

> - Density - ie how densely connected are the banks? Pre-Lehman = very dense
> - Clustering (transitivity) - how interwoven groups in the network are
> - Path length - linkedin reference
> - Degree 

```{r echo=F, messages = F, include=F, cache=F, results = 'asis'}
library(igraph)
library(quantmod)
library(igraph)
library(stockchartR)
library(networkD3)
```

## *Non-Random* **Regular** Network 


```{r}
set.seed(7) 
e <- erdos.renyi.game(100, 0.1) #100 Nodes, Prob edge btwen two vertices = .10
plot(e)
```

##Using R to interact with visualizations to understand the SIFI
```{r echo=F}

# Load data
data(MisLinks)
data(MisNodes)

# Plot
forceNetwork(Links = MisLinks, Nodes = MisNodes,
            Source = "source", Target = "target",
            Value = "value", NodeID = "name",
            Group = "group", opacity = 0.8)
```

## Properties of this Regular Non-Random Network

Easy to calculate these in R;

```{r}
graph.density(e) # Ratio of no. of edges vs no. possible edges
transitivity(e) # Prob that adjacent vertices are connected
average.path.length(e) # Average dist between vertices
```

## Obj 2: Identifying Systematically Important Financial Institutions (SIFIs)

> 1 There are some banks that are so important, that if they were to disappear the market itself would suffer greatly.

> 2 We can quantify that risk by defining some measure of "centrality"
    - Degree
    - Betweeness
    - Closeness
    - Eigenvector
    
##Obj 3: Let's create an Index value, SIFI = Index > 2.5

```{r IndexValue, echo=F}

data <- read.csv2("~/Data_Science/FinanceMemos/networktable.csv", stringsAsFactors=FALSE)
g <- graph.data.frame(data)
degree <- degree(g, normalized = T)
between <- betweenness(g, normalized = T)
closeness <- closeness(g, normalized = T)
eigenv <- evcent(g, directed = T)$vector

normalizedata <- function(x) x / mean(x)
index <- (normalizedata(degree) + normalizedata(between) + normalizedata(closeness) + normalizedata(eigenv) )/ 4
summary(index)
```
```{r echo=F, fig.width=5, fig.height=4.5}
hist(index)
```


##Which banks do we watch? Set cutoff to 3.5

These are the banks (by number) that we will watch for downgrades by a debt rating agency due to rising leverage ratios. 

```{r}
names(which(index>3.5))
```

##What instrument do we use to go short?

When these banks reach the critical threshold in our model, we short the financials sector using **XLF**. library(c("stockchartR", "quantmod")
```{r chart, echo=F, warning=F, message = F, }
xlf <- getSymbols("xlf", auto.assign=FALSE, from="2013-01-01")

XLFchart <- stockchartR::stockchart(adjustOHLC(xlf))
XLFchart

```
