---
title: "Financial Networks"
author: "Scott Jacobs, Andrea Bernard, Justin Salvador, Haixia Zhang"
date: "January 30, 2016"
output: ioslides_presentation
---

## Understanding Financial Networks Informs Trading

Financial networks is a map of which bank has lent money to some other bank. Analyzing the network can help identify banks that pose inherent risks to the system.

> - Guide overall portfolio strategy
> - Inform clients of potential pitfalls
> - Identify specific bank stocks to short

## Properties of Financial Networks (igraph package)

> - Density 
> - Clustering (transitivity)
> - Path length 
> - Degree 
```{r echo=F, messages = F, include=F, cache=F, results = 'asis'}
library(igraph)
```

## *Non-Random* **Regular** Network


```{r}
set.seed(7) 
e <- erdos.renyi.game(100, 0.1) #100 Nodes, Prob edge btwen two vertices = .10
plot(e)
```

## Properties of this Regular Non-Random Network

```{r}
graph.density(e) # Ratio of no. of edges vs no. possible edges
transitivity(e) # Prob that adjacent vertices are connected
average.path.length(e) # Average dist between vertices
```
